---
title: "Astrological Analysis of the PWHL"
subtitle: "By someone who doesn't know anything about astrology or hockey"
author: "Alexandra McSween"
format:
  revealjs: 
    theme: sky
---

```{r}
#devtools::install_github(repo = "sportsdataverse/fastRhockey")

library(fastRhockey)
library(tidyverse)
library(DescTools)
library(janitor)
source("./fixed_functions.R")

colours <- list(
"BOS" = "#154734",
"MIN" = "#2E1A47",
"MON" = "#862633",
"NY" = "#00B2A9",
"OTT" = "#A6192E",
"TOR" = "#307FE2"
)




# create a nice function to lapply to our team vector

team_stats <- function() {
 
  # get vector of team abbreviations
  teams <- fastRhockey::pwhl_teams()$team_code

  all_teams <- data.frame()

  for (team in teams){

  # here we use our modified functions
  df_stats <- pwhl_stats_fix(position = "skater", team = team, season = 2025) |>
    mutate(team = case_when(
      team == "MTL" ~ "MON",
      .default = team
      )
    )

  df_team <- pwhl_team_roster_fix(
    team = team,
    season = 2025
  ) %>%
      mutate(
        sign = DescTools::Zodiac(as.Date(dob))
      ) %>%
      merge(
        df_stats,
        by.x = c("player_id","player_name", "team"),
        by.y = c("player_id", "player","team")
      )
    
    all_teams <- rbind(all_teams, df_team)
  }

  return(all_teams)

}

# get data

all_teams <- team_stats()

print(all_teams |> head())

```

## Astrological Composition of the PWHL

```{r}
# let's look at totals by sign

all_teams %>% ggplot(aes(sign, fill = team)) +
  geom_bar() + 
  scale_fill_manual(values = colours) +
  labs(
    title = "Signs of the PWHL by team",
    fill = "Team"
  ) +
  theme(
    plot.title = element_text(size=24)
  )

```

## Astrological Composition of the PWHL

```{r}
# the above is a little overwhelming, let's convert to type of sign

all_teams <- all_teams %>% mutate(
  type = case_when(
    sign %in% c("Cancer", "Scorpio", "Pisces") ~ "Water", 
    sign %in% c("Aries", "Leo", "Sagittarius") ~ "Fire", 
    sign %in% c("Taurus", "Virgo", "Capricorn") ~ "Earth",
    sign %in% c("Gemini", "Libra", "Aquarius") ~ "Air"
  )
)

all_teams %>% ggplot(aes(type, fill = team)) +
  geom_bar() + 
  scale_fill_manual(values = colours) +
  labs(
    title = "Sign types of the PWHL by team",
    fill = "Team"
  ) +
  theme(
    plot.title = element_text(size=24)
)

```

## Astrological Composition of the PWHL

```{r}

all_teams %>% 
  ggplot(aes(sign, fill = position.x)) +
  geom_bar() +
  labs(
    title = "Signs of the PWHL by position",
    fill = "Position"
  ) +
  theme(
    plot.title = element_text(size=24)
)


```

## Analysis

Let's explore the astrological composition of our top scorers and top penalty minuters!

## Most goals

```{r}

# Grab the top 10 scorers

tail(all_teams[order(as.numeric(all_teams$goals)),], 10) %>%
  ggplot(aes(sign, fill = team)) +
  geom_bar() +
  scale_fill_manual(values = colours) +
  labs(
    title = "Sign of top 10 scorers",
    fill = "Team"
  ) +
  theme(
    plot.title = element_text(size=24),
    text = element_text(family = "Quicksand")
    )

```


## All goals by sign

```{r}

all_teams %>% group_by(team, sign) %>%
  summarize(goals = sum(as.numeric(goals))) %>%
  ggplot(aes(sign, goals, fill = team)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = colours) +
    labs(
    title = "Goals by sign",
    fill = "Team"
  ) +
    theme(
      plot.title = element_text(size=24)
    )
```

## All goals by sign

```{r}
# this could be skewed if there are many libras in the PWHL, for example.

all_teams %>% group_by(team, sign) %>%
  summarize(goals = sum(as.numeric(goals)), n = n()) %>%
  mutate(adjusted_goals = goals / n) %>%
  ggplot(aes(sign, adjusted_goals, fill = team)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = colours) +
    labs(
    title = "Proportional Goals by sign",
    fill = "Team"
  ) +
    theme(
      plot.title = element_text(size=24)
    )

```

## Penalty minutes
```{r}

tail(all_teams[order(as.numeric(all_teams$penalty_minutes)),], 10) %>%
  ggplot(aes(sign, fill = team)) +
  geom_bar() + 
  scale_fill_manual(values = colours) +
  labs(
    title = "Sign of top 10 penalty-havers",
    fill = "Team"
  ) +
  theme(
    plot.title = element_text(size=24)
  )
```

## All penalty minutes by sign

```{r}
# sim for penalty min

all_teams %>% group_by(team, sign) %>%
  summarize(penalty_minutes = sum(as.numeric(penalty_minutes))) %>%
  ggplot(aes(sign, penalty_minutes, fill = team)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = colours) +
    labs(
    title = "Penalty minutes by sign",
    fill = "Team"
  ) +
    theme(
      plot.title = element_text(size=24)
    )
```

## All penalty minutes by sign

```{r}

all_teams %>% group_by(team, sign) %>%
  summarize(penalty_minutes = sum(as.numeric(penalty_minutes)), n = n()) %>%
  mutate(adjusted_penalty_minutes = penalty_minutes / n) %>%
  ggplot(aes(sign, adjusted_penalty_minutes, fill = team)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = colours) +
    labs(
    title = "Proportional penalty minutes by sign",
    fill = "Team"
  ) +
    theme(
      plot.title = element_text(size=24)
    )
```

## Don't forget the goalies!!

```{r}

goalies <- pwhl_stats() %>% 
  merge(
    all_teams, on = "player_name"
  )

surnames <- unlist(strsplit(goalies$player_name, " "))

surnames <- surnames[c(FALSE, TRUE)]

goalies %>%
  ggplot(
  aes(player_name, save_percentage, fill = sign)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Save Percentage by Goalie (and her sign)",
    fill = "Sign",
    x = "Player",
    y = "Save Percentage"
  ) +
  scale_x_discrete(labels = surnames) +
  theme(
    plot.title = element_text(size=24)
  )


```

## Acknowledgements

Thank you to Brittny Vongdara for sharing her idea to analyze PWHL data!

Thank you to Murdoch McSween for sharing his astrology expertise!
